---
title: "Unit 11_Time Series"
author: "John Olanipekun"
date: "11/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
library(e1071)
```


```{r}
library(fpp)
```


# 1. SES MODEL FOR AUS AIR
The objective is to capture the trend in the data and predict out 3 years.

```{r}
data(ausair)
#returns a ts object for just that section of the historical data
air = window(ausair, start = 1990, end = 2004)

# Always plot the data first! 
plot(air,ylab = "Airline Passegners", xlab = "Year", main = "Airline Passengers")


```

### the ses model
```{r}
#fit 3 different simple exponential smoothing models ... how are they different?
# what does the h parameter do? 
fit1 = ses(air, initial = "simple",alpha = .2,h = 3) #alpha is the smoothing parameter,simple =1st value of the data set, h =predict out 3 years or 3 time points.
fit2 = ses(air,initial = "simple",alpha = .6, h = 3)
fit3 = ses(air, h = 3) #defaults

#plots models on top of the original data



```



```{r}
# the forecast package has a nice accuracy funciton with various metrics just pass it the 
# the model and the data!  (This is the "training" data)
accuracy(fit1, ausair)
accuracy(fit2, ausair)
accuracy(fit3, ausair)


```



```{r}
#Reset the plot
plot(air,ylab = "Airline Passegners", xlab = "Year", type = "o", xlim = c(1990, 2008),ylim = c(15,50), main = "Airline Passengers")
#Plot the estimated values from the models .. the "fitted" values are the training values.
lines(fitted(fit1), col = "blue", type = "o") #alpha = 0.2
lines(fitted(fit2), col = "red", type = "o") #alpha = 0.6
lines(fitted(fit3), col = "green", type = "o") #default params, alpha = 0.8

```



```{r}
# let s now take out the  $mean values are the forecasts.
#In SES the forecast is the mean of the previous value
lines(fit1$mean, col = "blue", type = "o")
lines(fit2$mean, col = "red", type = "o")
lines(fit3$mean, col = "green", type = "o") #looks better as it follows the original data
```


```{r}
# These are the actual values!  Compare visually with the forecasts! Not very great.
air2008 = window(ausair, start = 1990, end = 2007)
points(air2008, type = "o")

```


```{r}
# Compare the forecasts with the actual values with various fit metrics.  
accuracy(fit1, air2008)
accuracy(fit2, air2008)
accuracy(fit3, air2008)
```


###Holts
#2 Holt's Linear Trend Model for AUS AIR: Additive and multiplicative models
Multiplicative model gives a constant GROWTH RATE for forecast.
Additive model gives a constant RATE for forecast..

We need to later apply damping to make the predicted trend more realistic and not get that constant rate. we want the rate to flatten out over time.

```{r}
#Holts needs two values: an initial value of the series and value for initial change in the series(i.e. slope) (can be simply y2 - y1)
#(or fit a linear trend model to the first few observations and use resulting estimate of intercept and slope parameters as initial value for level and trend respectively)
#alpha (level of series smoothing) and beta (slope smoothing) are smoothing params. h means predict out 5(5 years here).
fit1h = holt(air, alpha = .8, beta = .2, initial = "simple", h = 5) #linear trend (additive model)
fit1h$model$state #l = level, b = slope for each state in the model.
fit2h = holt(air, alpha = .8, beta = .2, initial = "simple", exponential = TRUE, h = 5)#exponential means I am assuming a multiplicative model rather than additive.

```



```{r}
# Check out estimated values of the "training" data from the first holt model 
fitted(fit1h) #output includes the predicted values.
# Check out the forecast value (h of them)
fit1h$mean #the 5 predicted values

```

###plot the fitted values out on original data plot
```{r}
# Reset the Plot!
plot(air,ylab = "Airline Passegners", xlab = "Year", type = "o", xlim = c(1990, 2009),ylim = c(15,60))
#Plot each models estimated values of the training data (Do these one by one to see the differences)
lines(fitted(fit1h),col = "blue", type= "o") #additive model
lines(fitted(fit2h), col = "red", type= "o")#multiplicative effect model(exponential)
#Plot each models forecasts (Do these one by one to see the differences)
lines(fit1h$mean, col = "blue", type= "o") #just a straight a line
lines(fit2h$mean,col = "red", type= "o")#exponential is more optimistic as it goes up and curves accordingly

#predicted values list
fit1h$mean
fit2h$mean

```

###Damping
```{r}
# Fit another model ... damped!  
fit3h = holt(air, alpha = .8, beta = .2, damped = TRUE, initial = "optimal", h = 5)
fit3h.init = holt(air, alpha = .8, beta = .2, damped = TRUE, initial = "simple", h = 5)#use initial value of series
# Plot the fitted value (estimated from training data)
lines(fitted(fit3h), col = "darkgreen", type= "o")
# Plot the forecasts
lines(fit3h$mean,col = "darkgreen", type= "o")#damped trend, more realistic. trend reduces as we go out in time

lines(fitted(fit3h.init),col = "yellow", type= "o") #result is the same as "optimal"

```

###Evaluation
Using the fit$model method you can compare the stats for the model. lower AIC, BIC s are better model. Holt damped is better than other in this case.
Accuracy. Calculated and compared.
```{r}
# Fit another model ... what is the difference?  
fit4h = holt(air, alpha = .8, beta = .2, damped = TRUE, initial = "optimal", exponential = TRUE, h = 5)
# Plot the fitted value (estimated from training data)
#best fit of them all (damped multiplicative method)
lines(fitted(fit4h), col = "cyan", type= "o")
#Plot the forecasts
lines(fit4h$mean,col = "cyan", type= "o")

```


Accuracy. Calculated and compared.
```{r}
# with implicit Test set... it figures out by the time which are training and which are test. 
accuracy(fit1h, ausair)
accuracy(fit2h, ausair)
accuracy(fit3h, ausair)

```


```{r}
#with explicit Test set ... (same output)
airTest = window(ausair, start = 2005)
accuracy(fit1h, airTest)
accuracy(fit2h, airTest)
accuracy(fit3h, airTest)

```

Compare models in terms of slope and levels over time
```{r}
#plot level and slope components for Holt's linear trend
plot(fit1h$model$state)
plot(fit2h$model$state)
plot(fit3h$model$state) #exponential
```


```{r}
#Add the actual values to visually compare forecasts to actual values
air2008 = window(ausair, start = 1990, end = 2009)
points(air2008, type = "o")

```




Seasonal trend
Holts-Winter model

```{r}
#Load the data
data("austourists")


# returns a ts object.  
aust = window(austourists,start = 1999, end = 2004)

#fit an additive and multiplicative model
fit1s = hw(aust,seasonal = "additive",h = 40)
fit2s = hw(aust,seasonal = "multiplicative",h = 40)

#Plot the original data
plot(aust,ylab = "Australian Tourists", xlab = "Year", type = "o", xlim = c(1999, 2014),ylim = c(15,60))
#add the fitted values from the model (of the training data)
lines(fitted(fit1s),col = "blue", type= "o")
lines(fitted(fit2s), col = "red", type= "o")

#Now add the forecasts (add these one at a time)
lines(fit1s$mean, col = "blue", type= "o")
lines(fit2s$mean,col = "red", type= "o")

#Compare the accuracy
accuracy(fit1s,austourists)
accuracy(fit2s,austourists)

#add the actual values to visually compare the forecasts to the actual values. 
points(austourists, type = "o")

```


